---
title: "Time series features"
execute:
  echo: false
format: html
---

```{r}
#| label: setup
#| include: false
library(fpp3)
library(tidyverse)

ambulance_demand_tsb <- read_rds("../data/ambulance_demand_tsb.rds")
```

# Lab 3: time series features

## Ambulance demand data

This is super useful when you deal with many time series.

For the daily admissions time series with keys (with gender, injury), extract the strength of trend and seasonality

  *Do you see any useful insight?*

```{r}
#| label: ed-features
ambulance_demand_features <- ambulance_demand_tsb |>
  features(demand,feature_set(pkgs = "feasts"))
```

1. create  a scatterplot to show the strength of trend and seasonality features: 

```{r}
#| label: trend-seasonality-ambulance
ggplot(data = ambulance_demand_features, mapping = aes(spectral_entropy))+
  geom_histogram()
```

2. Using a feature that shows how difficult/easy is to forecast a time series and create a histogram to see the distribution of forecastability of the PBS data:

```{r}
#| label: entropy-ambulance
ggplot(data = ambulance_demand_features, mapping = aes(x = trend_strength, y = seasonal_strength_week))+
  geom_point()
```


## Pharmaceutical sales data

Let's know look at a bigger dataset.We explore the sales (demand) for Pharmaceutical products in Australia:

```{r}
#| label: PBS-features
## Compute features
PBS_feat <- PBS |>
  ????(Cost, feature_set(pkgs = "feasts")) |>
  na.omit()

```
 

Write r chunks to answer the following questions:

1. create  a scatterplot to show the strength of trend and seasonality features: 

```{r}
#| label: trend-seasonality
ggplot(data = PBS_feat, )
```

2. Using a feature that shows how difficult/easy is to forecast a time series and create a histogram to see the distribution of forecastability of the PBS data:

```{r}
#| label: entropy
ggplot(data = PBS_feat, )
```


we can only look at the handful number of features using visualizations, you can also use dimensional reduction techniques such as Principal Component Analysis:

Use a feature-based approach to look for outlying series in `PBS` data set. use ? PBS to know more about PBS data.

What is unusual about the series you identify as outliers?

```{r}
#| label: pca
library(broom)

## Compute principal components
PBS_prcomp <- PBS_feat |>
  select(-Concession, -Type, -ATC1, -ATC2) |>
  prcomp(scale = TRUE) |>
  augment(PBS_feat)

## Plot the first two components
PBS_prcomp |>
  ggplot(aes(x = ???, y = ???)) +
  geom_point()

## Pull out most unusual series from first principal component
outliers <- PBS_prcomp |>
  filter(.fittedPC1 > 7)
outliers

## Visualise the unusual series
PBS |>
  semi_join(outliers, by = c("Concession", "Type", "ATC1", "ATC2")) |>
  autoplot(Cost) +
  facet_grid(vars(Concession, Type, ATC1, ATC2)) +
  labs(title = "Outlying time series in PC space")
```
